R Notebook

### 2022-12-06 FEM_R/Loneliness_Social_Isolation/PAPER_WORKING_SCRIPT.Rmd
Added some functions for calculating Life Years and DFLYs from summary output. Ran these comparisons on baseline and intervention cases, not the outcome we wanted...

```{r}
# Directories
base_output_dir <- "output/COMPLETE/"
input_dir <- "input_data/"

@@

# Inputs
long <- read_dta(paste0(input_dir, 'ELSA_long.dta'))
stock <- read_dta(paste0(input_dir, 'ELSA_stock.dta'))
repl <- read_dta(paste0(input_dir, 'ELSA_repl.dta'))
trans <- read_dta(paste0(input_dir, 'ELSA_transition.dta'))

# Outputs
base_full <- read_dta(paste0(base_output_dir, 'ELSA_core_base/ELSA_core_base_summary.dta'))
cohort_full <- read_dta(paste0(base_output_dir, 'ELSA_core_cohort/ELSA_core_cohort_summary.dta'))
```

@@

@@

@@

### Life Years

@@

>>>>

```{r}
## Directories
base_output_dir <- "output/LNLY_SOCISO/"
input_dir <- "input_data/"

@@

## Inputs
long <- read_dta(paste0(input_dir, 'ELSA_long.dta'))
stock <- read_dta(paste0(input_dir, 'ELSA_stock.dta'))
repl <- read_dta(paste0(input_dir, 'ELSA_repl.dta'))
trans <- read_dta(paste0(input_dir, 'ELSA_transition.dta'))

## Outputs
### Baseline
base_full <- read_dta(paste0(base_output_dir, 'ELSA_core_base/ELSA_core_base_summary.dta'))
cohort_full <- read_dta(paste0(base_output_dir, 'ELSA_core_cohort/ELSA_core_cohort_summary.dta'))
### Interventions
no.lonely <- read_dta(paste0(base_output_dir, 'ELSA_nolnly_cohort/ELSA_nolnly_cohort_summary.dta'))
no.sociso <- read_dta(paste0(base_output_dir, 'ELSA_nosociso_cohort/ELSA_nosociso_cohort_summary.dta'))
no.lnly.sociso <- read_dta(paste0(base_output_dir, 'ELSA_nolnly_nosociso_cohort/ELSA_nolnly_nosociso_cohort_summary.dta'))
```

@@

Looking at three outcomes in particular, and maybe some other supporting stuff if its interesting enough.

The three outcomes are:
- Life Years
- Disability-free Life Years
- Disease-free Life Years

@@

### Cohort

Can work out Life Years and DFLYs from summary output for cohort populations.

We do this by multiplying the number of people alive by 2 each year to get the total number of people living (life years) per wave (2 years each wave). We then get the total number of life years for all people at each wave, and can therefore work out the average life years per person by dividing the total by number of people alive at each wave.

Using these calculation we can make comparisons between baseline and intervention scenarios in terms of average life years.

```{r "LifeYears"}
# Calculate average life years for an entire cohort from 50
cohortLifeyear <- function(c) {
  c.tot.lfyrs <- c$m_endpop_all * 2
  c.sum.tot.lfyrs <- sum(c.tot.lfyrs)
  c.avg.lfyr <- c.sum.tot.lfyrs / c$m_endpop_all[1]
  return(c.avg.lfyr)
}

# Compare cohort LifeYears from 50 between two summary outputs (baseline vs int)
lifeyearComparison <- function(c, i, int.name) {

  c.avg.lfyr <- cohortLifeyear(c)

  i.avg.lfyr <- cohortLifeyear(i)

  print(int.name)
  print(paste('Cohort Average Lifeyears:        ', signif(c.avg.lfyr, digits=4)))
  print(paste('Intervention Average Lifeyears:  ', signif(i.avg.lfyr, digits=4)))
  print(paste('Increase:                        ', signif(x = (i.avg.lfyr - c.avg.lfyr), digits=4)))
}

# This function calculates the average number of life years in the model when diagnosed with a chronic disease
diseaseVarLifeyear <- function(c, var) {

  prevVar <- paste0('p_', var, '_all')

  c.tot.var.lfyrs <- c[,prevVar] * c$m_endpop_all * 2
  c.sum.var.lfyrs <- sum(c.tot.var.lfyrs)
  c.avg.var.lfyrs <- c.sum.var.lfyrs / c$m_endpop_all[1]

  print(paste0(var, ':    ', signif(c.avg.var.lfyrs, digits=4)))
}
```

```{r "Disability"}
# Calculate disability life years
disabilityLifeYearsCohort <- function(c) {
  total_dis <- c$a_disabled_all * c$m_endpop_all
  sum_tot_dis <- sum(total_dis)
  avg_dis <- sum_tot_dis / c$m_endpop_all[1]
  return(avg_dis)
}

# Compare disability life years between scenarios
disabilityLifeYearComp <- function(c, i, int.name) {
  c.avg.dis <- disabilityLifeYearsCohort(c)
  i.avg.dis <- disabilityLifeYearsCohort(i)
  print(int.name)
  print(paste('Cohort Average Disabled Lifeyears:        ', signif(c.avg.dis, digits=4)))
  print(paste('Intervention Average Disabled Lifeyears:  ', signif(i.avg.dis, digits=4)))
  print(paste('Increase:                        ', signif(x = (i.avg.dis - c.avg.dis), digits=4)))
}

# Calculate disability free life years
disabilityFreeLifeYearsCohort <- function(c) {
  total_disfree <- c$a_not_disabled_all * c$m_endpop_all
  sum_tot_disfree <- sum(total_disfree)
  avg_disfree <- sum_tot_disfree / c$m_endpop_all[1]
  return(avg_disfree)
}

# Compare disability free life years between scenarios
disabilityFreeLifeYearsComp <- function(c, i, int.name) {
  c.avg.dis <- disabilityFreeLifeYearsCohort(c)
  i.avg.dis <- disabilityFreeLifeYearsCohort(i)
  print(int.name)
  print(paste('Cohort Average Disability Free Lifeyears:        ', signif(c.avg.dis, digits=4)))
  print(paste('Intervention Average Disability Free Lifeyears:  ', signif(i.avg.dis, digits=4)))
  print(paste('Increase:                                        ', signif(x = (i.avg.dis - c.avg.dis), digits=4)))
}
```

```{r "Disease"}
# Calculate disease life years
diseaseLifeYearsCohort <- function(c) {
  total_dis <- c$a_anydisease_all * c$m_endpop_all
  sum_tot_dis <- sum(total_dis)
  avg_dis <- sum_tot_dis / c$m_endpop_all[1]
  return(avg_dis)
}

# Compare disease life years between scenarios
diseaseLifeYearComp <- function(c, i, int.name) {
  c.avg.dis <- diseaseLifeYearsCohort(c)
  i.avg.dis <- diseaseLifeYearsCohort(i)
  print(int.name)
  print(paste('Cohort Average Diseased Lifeyears:        ', signif(c.avg.dis, digits=4)))
  print(paste('Intervention Average Diseased Lifeyears:  ', signif(i.avg.dis, digits=4)))
  print(paste('Increase:                        ', signif(x = (i.avg.dis - c.avg.dis), digits=4)))
}

# Calculate disease free life years
diseaseFreeLifeYearsCohort <- function(c) {
  total_disfree <- c$a_nodisease_all * c$m_endpop_all
  sum_tot_disfree <- sum(total_disfree)
  avg_disfree <- sum_tot_disfree / c$m_endpop_all[1]
  return(avg_disfree)
}

# Compare disease free life years between scenarios
diseaseFreeLifeYearsComp <- function(c, i, int.name) {
  c.avg.dis <- diseaseFreeLifeYearsCohort(c)
  i.avg.dis <- diseaseFreeLifeYearsCohort(i)
  print(int.name)
  print(paste('Cohort Average Disease Free Lifeyears:        ', signif(c.avg.dis, digits=4)))
  print(paste('Intervention Average Disease Free Lifeyears:  ', signif(i.avg.dis, digits=4)))
  print(paste('Increase:                                        ', signif(x = (i.avg.dis - c.avg.dis), digits=4)))
}
```

### Whole Population

@@

### Life Years
```{r}
coh.ly <- cohortLifeyear(cohort_full)

# Total lifeyears
print(paste('Cohort Average Lifeyears:        ', signif(coh.ly, digits=4)))

# Cohort disease life years
diseaseVarLifeyear(cohort_full, 'cancre')
diseaseVarLifeyear(cohort_full, 'diabe')
diseaseVarLifeyear(cohort_full, 'hearte')
diseaseVarLifeyear(cohort_full, 'lunge')
diseaseVarLifeyear(cohort_full, 'hibpe')
diseaseVarLifeyear(cohort_full, 'stroke')
diseaseVarLifeyear(cohort_full, 'alzhe')
diseaseVarLifeyear(cohort_full, 'demene')
```

@@

### Burden of Loneliness
```{r}
coh.ly <- cohortLifeyear(cohort_full)
nolnly.ly <- cohortLifeyear(no.lonely)

lifeyearComparison(cohort_full, no.lonely, 'No Loneliness')
lifeyearComparison(cohort_full, no.sociso, 'No Social Isolation')
lifeyearComparison(cohort_full, no.lnly.sociso, 'No Loneliness OR Social Isolation')
```

### 2022-12-06 FEM_R/calculate_dflys_from_summary.Rmd
Added some functions for calculating Life Years and DFLYs from summary output. Ran these comparisons on baseline and intervention cases, not the outcome we wanted...

# Setup
```{r "setup", include=FALSE}
# Need to set up the environment
workingDir <- "/home/luke/Documents/E_FEM_clean/tmp_output/"
knitr::opts_knit$set(root.dir = workingDir)

@@

# Hearte (cohort and pop)
heart_c <- read_dta('ELSA_core_remove_hearte_c/ELSA_core_remove_hearte_c_summary.dta')
#heart_p <- read_dta('ELSA_core_remove_hearte_p/ELSA_core_remove_hearte_p_summary.dta')

# Smoking (cohort)
smok <- read_dta('ELSA_core_remove_smoken/ELSA_core_remove_smoken_summary.dta')

# BMI (cohort)
#nomorbid <- read_dta('ELSA_core_nomorbid/ELSA_core_nomorbid_summary.dta')
#obese2 <- read_dta('ELSA_core_noobese2/ELSA_core_noobese2_summary.dta')
obese1 <- read_dta('ELSA_core_noobese1/ELSA_core_noobese1_summary.dta')

# Drinking (pop)
nodrink_p <- read_dta('ELSA_core_remove_drink_p/ELSA_core_remove_drink_p_summary.dta')
nodrink_c <- read_dta('ELSA_core_remove_drink_c/ELSA_core_remove_drink_c_summary.dta')
#noBadDrink <- read_dta('ELSA_core_remove_problem_drink_p/ELSA_core_remove_problem_drink_p_summary.dta')
```

@@

### Cohort
```{r}
# Cohort life years
c <- cohort
c.ly <- cohortLifeyear(c)
print(paste0('Cohort Life Years: ', signif(c.ly, digits=4)))
print('-------------------')

@@

>>>>

# Setup
```{r "setup", include=FALSE}
# Need to set up the environment
workingDir <- "/home/luke/Documents/E_FEM_clean/E_FEM/output/COMPLETE/"
knitr::opts_knit$set(root.dir = workingDir)

@@

# Hearte (cohort and pop)
#heart_c <- read_dta('ELSA_core_remove_hearte_c/ELSA_core_remove_hearte_c_summary.dta')
#heart_p <- read_dta('ELSA_core_remove_hearte_p/ELSA_core_remove_hearte_p_summary.dta')

# Smoking (cohort)
#smok <- read_dta('ELSA_core_remove_smoken/ELSA_core_remove_smoken_summary.dta')

# BMI (cohort)
#nomorbid <- read_dta('ELSA_core_nomorbid/ELSA_core_nomorbid_summary.dta')
#obese2 <- read_dta('ELSA_core_noobese2/ELSA_core_noobese2_summary.dta')
#obese1 <- read_dta('ELSA_core_noobese1/ELSA_core_noobese1_summary.dta')

# Drinking (pop)
#nodrink_p <- read_dta('ELSA_core_remove_drink_p/ELSA_core_remove_drink_p_summary.dta')
#nodrink_c <- read_dta('ELSA_core_remove_drink_c/ELSA_core_remove_drink_c_summary.dta')
#noBadDrink <- read_dta('ELSA_core_remove_problem_drink_p/ELSA_core_remove_problem_drink_p_summary.dta')
```

@@

### Cohort
```{r}
# Cohort life years
c <- cohort_full
c.ly <- cohortLifeyear(c)
print(paste0('Cohort Life Years: ', signif(c.ly, digits=4)))
print('-------------------')

@@

```{r}
